Multidimensional Persistence Module Classification via Lattice-Theoretic Convolutions

### 2020-10-10 paper/paper.tex
minor edits

Multidimensional persistence generalizes single-dimensional persistent homology
in order to tackle filtrations parameterized in multiple dimensions.
% Such
% filtrations arise, for example, in a Rips filtration of a finite metric space
% for which the finite metric space itself carries a filter function. %is parameterized by a signal
% %attributing a vector in $\R^d$ to every point of the metric space.
% Restricting
% to points in the metric space appearing below a certain coordinate-wise
% threshold results in a multi-filtration of the corresponding Rips complexes.
%Multidimensional persistence boasts stability properties similar to those in one
%dimension~\cite{lesnick_theory_2015}.
Unfortunately, there is no complete
compact barcode-like characterization of multidimensional persistence
modules~\cite{carlsson_theory_2009}. We must make do with incomplete invariants.
There are various algebraic invariants; in this paper we will use the
\textit{Hilbert function} and the \textit{multi-graded Betti numbers}, both of
which are $\mathbb{N}$-valued functions on the parameter space.
%Their names arise from their relationship with the homological algebra of
%persistence modules.
The Hilbert function is nothing more than the pointwise
(topological) Betti numbers\footnote{\textit{Caveat lector:} the multi-graded
Betti numbers are not the same as the topological Betti numbers
$\beta_i(\mathbb{X}) = \dim(H_i(\mathbb{X}))$.}, and the multi-graded Betti
numbers have an interpretation in terms of births and deaths~\cite{knudson_refinement_2008}.

@@

\section{Experiments}
We use a small portion of the Princeton ModelNet
dataset~\cite{zhirong_wu_3d_2015} as a source of finite
metric spaces. This dataset consists of hundreds of 3-dimensional CAD models
representing objects from 40 classes. We select two of the classes and sample
points from the 3D models to produce finite metric spaces embedded in $\R^3$.
We then compute the corresponding multidimensional persistence modules, from
which we produce features used as an input to a convolutional neural net
classifier.

The pipline thus begins with a 3D model, which is sampled to produce a point
cloud in $\R^3$. This point cloud then produces a bifiltered simplicial complex,
whose persistent homology we calculate using
RIVET~\cite{lesnick_interactive_2015}, sampled at a discrete grid of $40 \times
40$ points, producing lattice-indexed signals given by the Hilbert function and
the multi-graded Betti numbers. These are then passed to the classifier, which
produces a class prediction.

As the filter function on these data sets, we use
\[\rho_{\text{codense}}(x;k) = \frac{k \diam(\mathcal M)}{\sum_{y \in N_k(x)}
    d(x,y)},\]
where $N_k(x)$ is the set of the $k$ nearest neighbors to $x$. This is the
\introduce{codensity} filtration, so named because the points in the densest
regions of $\mathcal M$ will appear first.

>>>>

Multidimensional persistence generalizes single-dimensional persistent homology
in order to tackle filtrations parameterized in multiple dimensions.
% Such
% filtrations arise, for example, in a Rips filtration of a finite metric space
% for which the finite metric space itself carries a filter function. %is parameterized by a signal
% %attributing a vector in $\R^d$ to every point of the metric space.
% Restricting
% to points in the metric space appearing below a certain coordinate-wise
% threshold results in a multi-filtration of the corresponding Rips complexes.
%Multidimensional persistence boasts stability properties similar to those in one
%dimension~\cite{lesnick_theory_2015}.
Unfortunately, there is no complete
compact barcode-like characterization of multidimensional persistence
modules~\cite{carlsson_theory_2009}. We must make do with incomplete invariants.
There are various algebraic invariants; in this paper we will use the
\textit{Hilbert function} and the \textit{multi-graded Betti numbers}, both of
which are $\mathbb{N}$-valued functions on the parameter space.
%Their names arise from their relationship with the homological algebra of
%persistence modules.
The Hilbert function is nothing more than the pointwise
(topological) Betti numbers\footnote{\textit{Caveat lector:} the multi-graded
Betti numbers are not the same as the topological Betti numbers
$\beta_i(\mathbb{X}) = \dim(H_i(\mathbb{X}))$.}, and the multi-graded Betti
numbers have a geometric interpretation in terms of births and deaths~\cite{knudson_refinement_2008}.

@@

\section{Experiments}
We use a small portion of the Princeton ModelNet
dataset~\cite{zhirong_wu_3d_2015} as a source of finite
metric spaces. This dataset consists of hundreds of 3-dimensional CAD models
representing objects from 10 classes. We select two of the classes and sample
points from the 3D models to produce finite metric spaces embedded in $\R^3$.
We then compute the corresponding multidimensional persistence modules, from
which we produce features used as an input to a convolutional neural net
classifier.

The pipline thus begins with a 3D polyhedral model, of which $3000$ vertices are sampled to produce a point
cloud in $\R^3$. This point cloud then produces a bifiltered simplicial complex,
whose persistent homology we calculate using
RIVET~\cite{lesnick_interactive_2015}, sampled at a discrete grid of $40 \times
40$ points, producing lattice-indexed signals given by the Hilbert function and
the multi-graded Betti numbers $\xi_0, \xi_1, \xi_2$--four features in total. These are then passed to the classifier, which
produces a class prediction.

As the filter function on these data sets, we use
\[\rho_{\text{codense}}(x;k) = \frac{1}{\frac{1}{k} \sum_{y \in N_k(x)}
    d(x,y)},\]
where $N_k(x)$ is the set of the $k$ nearest neighbors to $x$. This is the
\introduce{codensity} filtration, so named because the points in the densest
regions of $\mathcal M$ will appear first.
